A Generic and Efficient E-field Parallel Imaging Correlator for Next-Generation Radio Telescopes

### 2015-10-29 imaging_demo/ms.tex
Fix minor errors in previous version and resubmit
From f6809de (arxiv):
====================

In ms.tex,

(*) Mention calibration scales as number of calibrators as well as
number of antennas
(*) Correct the collecting area of HIRAX in the table
(*) Change caption of figure 9 to left and right instead of top and
bottom
(*) Update caption of figure 7

In a companion paper to appear soon, we demonstrate a novel calibration 
technique (EPICal) which leverages the data products formed by direct imaging 
correlators to estimate antenna complex gains. This method correlates the antenna 
electric field signals with an image pixel from the output of the correlator in 
the feedback calibration fashion outlined in \citealt{mor11} (illustrated in Fig. 
\ref{fig:MOFF-flowchart} by the arrow leading from the imager to the calibration 
block). Furthermore it allows for arbitrarily complex sky models, and following 
the MOFF algorithm places no restriction on array layout, and accounts for 
non-identical antenna beam patterns. Because only a single correlation is needed 
for each antenna, the computation complexity scales only as $N_{\mathrm{a}}$.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{MOFF_FX_comparison_10_random_source_positions_4_iterations_test_aperture_zoomed}
  % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_10_random_source_positions_4_iterations_test_aperture_zoomed.eps}
  \caption{Dirty images (top) and synthesized beams (bottom) obtained from 
    simulated data using antenna-based MOFF algorithm (left) and FX 
    visibility-based software holography (right). The solid black circles in the
    top panels indicate the simulated source positions. The antenna 
    auto-correlations at zero-spacing have been removed from the MOFF images. 
    The images in either case reconstruct the sources at the right locations 
    with the fluxes expected after multiplication by the antenna power pattern. 
    The synthesized beams from the two algorithms are well matched in size and
    shape. The overall modulation by the power pattern is seen clearly in both
    images.}
  \label{fig:MOFF-FX-image}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture}
  % \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture.eps}
  \caption{Map of difference between the synthesized beams obtained with the 
    two methods (top) and radial statistics of the synthesized beams and 
    their differences (bottom). The radial variations of rms of the synthesized 
    beam (gray) and that of the rms of the difference (black) are shown as 
    percentage of the peak synthesized beam. Radially averaged directional 
    antenna power response in absolute scale is shown in red and is to be read 
    with the scale on the right side of the axis. The maximum difference is of 
    the order of a few percent. Amplitude of the rms of the difference is 
    modulated by the power pattern of the antenna.}
  \label{fig:psf-diff}
\end{figure}

@@

\begin{table}
  \scriptsize
  \centering
  \caption{Radio telescopes and array layouts.}
  \label{tab:antenna-layouts}
  \begin{threeparttable}
  \begin{tabular}{ccccc} 
    \hline
    Telescope & Core & Number & Antenna & Frequency \\
              & size & of Antennas & size & \\
              & $b_\textrm{max}$ (in m) & $\Nant$ & $A_\textrm{a}$ (in m$^2$) & $f_0$ (in MHz) \\
    \hline
    MWA-112\tnote{a} & 1400 & 112 & 16 & 150 \\
    MWA-240\tnote{a} & 1400 & 240 & 16 & 150 \\
    MWA-496\tnote{a} & 1400 & 496 & 16 & 150 \\
    MWA-112\tnote{a} & 1400 & 1008 & 16 & 150 \\
    LOFAR-LC\tnote{b} & 3500 & 24 & 5809 & 50 \\
    LOFAR-HC\tnote{b} & 3500 & 48 & 745 & 150 \\
    LWA1 & 100 & 256 & 10 & 50 \\
    LWA-OV\tnote{c} & 200 & 256 & 10 & 50 \\
    HERA-19 & 70 & 19 & 154 & 150 \\
    HERA-37 & 98 & 37 & 154 & 150 \\
    HERA-331 & 294 & 331 & 154 & 150 \\
    HERA-6769\tnote{d} & 1330 & 6769 & 154 & 150 \\
    SKA1-LC\tnote{e} & 1000 & 750 & 962 & 150 \\
    SKA1-LCD\tnote{f} & 1000 & 192,000 & 2 & 150 \\
    CHIME & 100 & 1280 & 8 & 600 \\
    HIRAX\tnote{g} & 200 & 1024 & 6 & 600 \\
    \hline
  \end{tabular}
  \begin{tablenotes}
    \item[a] MWA-N denotes N tiles in the specified core diameter
    \item[b] LC and HC denotes low band and high band stations inside the 
      specified core diameter 
    \item[c] Owens Valley LWA
    \item[d] Hypothetically chosen to have a total collecting area of 
      1~km$^2$
    \item[e] This is the number of beamformed stations expected to be in the 
      core, roughly three-fourths of the total number
    \item[f] All dipoles inside the core are used as independent elements 
      without station beamforming
    \item[g] Hydrogen Intensity mapping and Real-time Analysis eXperiment
  \end{tablenotes}
  \end{threeparttable}
\end{table}

>>>>

In a companion paper to appear soon, we demonstrate a novel calibration 
technique (EPICal) which leverages the data products formed by direct imaging 
correlators to estimate antenna complex gains. This method correlates the antenna 
electric field signals with an image pixel from the output of the correlator in 
the feedback calibration fashion outlined in \citealt{mor11} (illustrated in Fig. 
\ref{fig:MOFF-flowchart} by the arrow leading from the imager to the calibration 
block). Furthermore it allows for arbitrarily complex sky models, and following 
the MOFF algorithm places no restriction on array layout, and accounts for 
non-identical antenna beam patterns. Because only a single correlation is needed 
for each antenna, the computation complexity scales only as $\sim N_{\mathrm{a}}N_\textrm{c}$, where $N_\textrm{c}$ is the number of calibrators in the field.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{MOFF_FX_comparison_10_random_source_positions_4_iterations_test_aperture_zoomed}
  % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_10_random_source_positions_4_iterations_test_aperture_zoomed.eps}
  \caption{Dirty images (top) and synthesized beams (bottom) obtained from 
    simulated data using EPIC implementation of antenna-based MOFF algorithm 
    (left) and visibility-based imaging (right). The solid black circles in the
    top panels indicate the simulated source positions. The antenna 
    auto-correlations at zero-spacing have been removed from the MOFF images. 
    The images in either case reconstruct the sources at the right locations 
    with the fluxes expected after multiplication by the antenna power pattern. 
    The synthesized beams from the two algorithms are well matched in size and
    shape. The overall modulation by the power pattern is seen clearly in both
    images.}
  \label{fig:MOFF-FX-image}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture}
  % \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture.eps}
  \caption{Map of difference between the synthesized beams obtained with the 
    two methods (left) and radial statistics of the synthesized beams and 
    their differences (right). The radial variations of rms of the synthesized 
    beam (gray) and that of the rms of the difference (black) are shown as 
    percentage of the peak synthesized beam. Radially averaged directional 
    antenna power response in absolute scale is shown in red and is to be read 
    with the scale on the right side of the axis. The maximum difference is of 
    the order of a few percent. Amplitude of the rms of the difference is 
    modulated by the power pattern of the antenna.}
  \label{fig:psf-diff}
\end{figure}

@@

\begin{table}
  \scriptsize
  \centering
  \caption{Radio telescopes and array layouts.}
  \label{tab:antenna-layouts}
  \begin{threeparttable}
  \begin{tabular}{ccccc} 
    \hline
    Telescope & Core & Number & Antenna & Frequency \\
              & size & of Antennas & size & \\
              & $b_\textrm{max}$ (in m) & $\Nant$ & $A_\textrm{a}$ (in m$^2$) & $f_0$ (in MHz) \\
    \hline
    MWA-112\tnote{a} & 1400 & 112 & 16 & 150 \\
    MWA-240\tnote{a} & 1400 & 240 & 16 & 150 \\
    MWA-496\tnote{a} & 1400 & 496 & 16 & 150 \\
    MWA-112\tnote{a} & 1400 & 1008 & 16 & 150 \\
    LOFAR-LC\tnote{b} & 3500 & 24 & 5809 & 50 \\
    LOFAR-HC\tnote{b} & 3500 & 48 & 745 & 150 \\
    LWA1 & 100 & 256 & 10 & 50 \\
    LWA-OV\tnote{c} & 200 & 256 & 10 & 50 \\
    HERA-19 & 70 & 19 & 154 & 150 \\
    HERA-37 & 98 & 37 & 154 & 150 \\
    HERA-331 & 294 & 331 & 154 & 150 \\
    HERA-6769\tnote{d} & 1330 & 6769 & 154 & 150 \\
    SKA1-LC\tnote{e} & 1000 & 750 & 962 & 150 \\
    SKA1-LCD\tnote{f} & 1000 & 192,000 & 2 & 150 \\
    CHIME & 100 & 1280 & 8 & 600 \\
    HIRAX\tnote{g} & 200 & 1024 & 28 & 600 \\
    \hline
  \end{tabular}
  \begin{tablenotes}
    \item[a] MWA-N denotes N tiles in the specified core diameter
    \item[b] LC and HC denotes low band and high band stations inside the 
      specified core diameter 
    \item[c] Owens Valley LWA
    \item[d] Hypothetically chosen to have a total collecting area of 
      1~km$^2$
    \item[e] This is the number of beamformed stations expected to be in the 
      core, roughly three-fourths of the total number
    \item[f] All dipoles inside the core are used as independent elements 
      without station beamforming
    \item[g] Hydrogen Intensity mapping and Real-time Analysis eXperiment
  \end{tablenotes}
  \end{threeparttable}
\end{table}
